---
title: "Clustering on simulated data"
author: "Zhiyuan Hu"
date: "10/03/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# BiocManager::install("Oshlack/splatter", dependencies = TRUE,
#          build_vignettes = TRUE,version = 1.7.2)
library(splatter)
library(scater)
library(SingleCellExperiment)
library(limma)
```

## Simulation dataset

```{r}
# sceset <- readRDS("../rds/20180725_allFT_Clincluster_12clusters_sceset.rds")
secretory <-  readRDS("../rds/20190120Fresh_secretory_9clusters_clincluster.rds")

secretory2 <- secretory[,secretory$clincluster_final %in% c("C3","C4","C6","C8","C10")]
```


```{r}
as.numeric(table(secretory$clincluster_final)/ncol(secretory))
```

```{r}
table(secretory$clincluster_final)
```


```{r}
dim(secretory)
```


```{r}
as.numeric(table(secretory$Patient2))
```

```{r}
params <- newSplatParams(nGenes = 15000, seed=234567,
                         group.prob = c(0.08,0.12,0.11,0.15,0.08, 0.2,  0.26),  
                         batchCells = as.numeric(table(secretory$Patient2)))
params
```


```{r}
# Simulate data using estimated parameters
sim <- splatSimulate(params, method = "groups")

sim <- normalize(sim)
plotPCA(sim, colour_by = "Group", shape_by = "Batch")
```


## Clustering

```{r}
source("clincluster/clincluster_functions.R")
sim <- PrepareData(sim, col.for.cluster = "Batch", do.scale = T)
```

#### High variable genes

```{r}
sim <- HighVarGenes(sim)
table(rowData(sim)$high.var)
```

The high variance genes

```{r}
ggplot(data = data.frame(gene.mean = rowData(sim)$gene.mean,
                         gene.dispersion = rowData(sim)$gene.dispersion,
                         high.var = rowData(sim)$high.var), 
       aes(x = gene.mean, y = gene.dispersion, col = high.var) ) +
      geom_point(alpha=0.4) 
```

#### Run tSNE by the log-transformed data

```{r}
set.seed(1234)
sim <- runTSNE(object = sim, ncomponents = 2, 
                     feature_set = rownames(sim)[rowData(sim)$high.var],
                     exprs_values = "logcounts", 
                     perplexity = min(50, floor(ncol(sim)/5)))

plotTSNE(sim, colour_by = "Group", shape = "Batch")
```

#### Calculate first 20 PCs

Calculate 20 PCs from high variance genes and the log-transformed counts.

```{r}
set.seed(34567)
sim <- runPCA(object = sim, ncomponents = 20, 
                    exprs_values = "logcounts", rand_seed = 34567,
                    feature_set = rownames(sim)[rowData(sim)$high.var == TRUE])

plotPCA(sim)
```


Plot the variance by PCs.

```{r}
plot(1:50, (attr(sim@reducedDims$PCA, "percentVar")[1:50])*100, pch = 20, xlab = "PC", ylab = "Standard Deviation of PC")
```

## Initial clustering

```{r}
set.seed(12345678)
sim <- InitialCluster(sim, k = 8, ncomponents = 1:11, n.neighbor = 6, spec.method = "kknn", iterations = 100)
```

#### Which clustering is better?

#### Visualisation of initial clusters

```{r, fig.height=7, fig.width=16}
p1 <- plotTSNE(sim[,sim$Batch == "Batch1"], colour_by = "initial.cluster")
p2 <- plotTSNE(sim[,sim$Batch == "Batch2"], colour_by = "initial.cluster")
p3 <- plotTSNE(sim[,sim$Batch == "Batch3"], colour_by = "initial.cluster")
p4 <- plotTSNE(sim[,sim$Batch == "Batch4"], colour_by = "initial.cluster")
p5 <- plotTSNE(sim[,sim$Batch == "Batch5"], colour_by = "initial.cluster")
cowplot::plot_grid(p1,p2,p3,p4,p5,nrow = 2)
```

```{r}
table(sim$Group, sim$initial.cluster)
```


## Limma

#### Remove low expressed genes

```{r}
matrix <- expm1(logcounts(sim))
keep <- rowSums(matrix > 1) > 5
sum(keep)
# 15508
```

```{r}
dge <- edgeR::DGEList(counts = matrix[keep,]) # make a edgeR object
rm(matrix,keep)
sim@colData$Patient2 <- as.factor(sim@colData$Batch)
design <- model.matrix(~  0 + initial.cluster, data = sim@colData)  # Use 0 because we do not need intercept for this linear model
design2 <- model.matrix(~  0 + Batch, data = sim@colData)  
colnames(design)
```

## Incoporate patients into contrast matrix

```{r}
v <- voom(dge, design, plot = TRUE)
```

#### Fit linear model for each gene given a series of arrays

```{r}
fit <- lmFit(v, design) # Linear Model for Series of Arrays

initial.clusters <- data.frame(id = colnames(design),
                               short_id = gsub(pattern = "initial.cluster", 
                                               replacement = "", x = colnames(design)),
                               patient = substr(colnames(design), start = 16, stop = 21))

## Automating makeContrasts call in limma
nc <- nrow(initial.clusters)
contrast_all <- gtools::permutations(v = as.character(initial.clusters$id), n = nc, r = 2)
contrast_all <- as.data.frame(contrast_all)
head(contrast_all)
```

#### Number of cells in each initial cluster

```{r}
initial.clusters$n_cells <- table(sim$initial.cluster)
n_cells_patients <- table(sim$Patient2)
initial.clusters$n_cells_patients <- n_cells_patients[match(initial.clusters$patient, names(n_cells_patients))]
initial.clusters$weight_cluster <- initial.clusters$n_cells/initial.clusters$n_cells_patients
initial.clusters$paste_weight_id <- paste(initial.clusters$id,"*",initial.clusters$weight_cluster, sep = "")

contrast_all$P1 <- substr(contrast_all$V1, start = 16, stop = 21) # patient 1
contrast_all$P2 <- substr(contrast_all$V2, start = 16, stop = 21) # patient 1
contrast_all$C1 <- NA
contrast_all$C2 <- NA
contrast_all$n_C1 <- NA
contrast_all$n_C2 <- NA
for(i in 1:nrow(contrast_all)) {
    contrast_all$C1[i] <- paste(initial.clusters$paste_weight_id[initial.clusters$patient == contrast_all$P1[i]], collapse = "+")
    contrast_all$C2[i] <- paste(initial.clusters$paste_weight_id[initial.clusters$patient == contrast_all$P2[i]], collapse = "+")
}
head(contrast_all)
```


```{r}
contrast_matrix <- apply(contrast_all, MARGIN = 1, function(x) return(paste(x[1],"-",x[2],"-(",x[5],")","+(", x[6],")", sep = "")))

cont.matrix <- makeContrasts(contrasts = contrast_matrix,levels=design)
cont.matrix[,5]
```


```{r}
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2)
```

#### Add weights to DEGs

The DE gene weight is decided by the fold change and the ratio of expression proportion.

```{r}
## parameter:
## logFC = 0.6
## p-value = 0.05
## weight = abs(logFC)*(expr_ratio1+0.01)/(expr_ratio2+0.01)
## expr_ratio_max > 0.25

n_deg2 <- matrix(0, ncol = nc, nrow = nc)  # number of DE genes
colnames(n_deg2) <- rownames(n_deg2) <- gsub(x = colnames(design)[1:nc], pattern = "initial.cluster",replacement = "")
for(i in 1:nc) {
    for(j in 1:nc) {
        if(i == j) {
            n_deg2[i,j] <- 0
        } else if (j < i) {
            coef_k = (i-1)*(nc-1)+j
        } else if (j > i) {
            coef_k = (i-1)*(nc-1)+j-1
        }
        
        if(i != j) {
            rls <- topTable(fit2, n = Inf, coef = coef_k, sort = "p", lfc = 0.6, p = 0.05 )
            if(nrow(rls) > 1) {
                v_expr <- logcounts(sim)[rownames(rls), sim$initial.cluster == rownames(n_deg2)[i]]
                rls$ratio1 <- rowSums(v_expr > 0.5)/ncol(v_expr)
                v_expr <- logcounts(sim)[rownames(rls), sim$initial.cluster == colnames(n_deg2)[j]]
                rls$ratio2 <- rowSums(v_expr > 0.5)/ncol(v_expr)
                rls$ratiomax <- rowMaxs(as.matrix(rls[,c("ratio1", "ratio2")]))
                rls$ratiomin <- rowMins(as.matrix(rls[,c("ratio1", "ratio2")]))
                rls <- rls[rls$ratiomax > 0.25, ]
                n_deg2[i,j] <- sum(apply(rls, MARGIN = 1, function(x) return(abs(x[1]) * (x[9]+0.01)/(x[10]+0.01)))) ## 0.01 is used here to exaggerate the differences of on-off genes
            } else if (nrow(rls) == 1) {
                n_deg2[i,j] <- sum(rls$logFC)
            }
            ## This eqaution take fold change and expression ratio into account
            ## Question: should we talk a upper limit to the weight?
        }
    }
}
```

## Final clustering

```{r}
## 7 clusters
hc <- hclust(as.dist(n_deg2))
hc.cluster <- cutree(hc, k = 7)

colData(sim)$clincluster <- hc.cluster[match(colData(sim)$initial.cluster, names(hc.cluster))]
colData(sim)$clincluster <- as.factor(colData(sim)$clincluster)
```

```{r}
## visualisation
hc <- hclust(as.dist(n_deg2))
plot(hc);rect.hclust(hc, k = 7, border = "red")
```


```{r}
df_plot <- as.data.frame(table(sim$clincluster, sim$Group))
colnames(df_plot) <- c("Cluster","Group","Freq")
df_plot$fill <- df_plot$Freq > 0
ggplot(df_plot, aes(x = Cluster, y = Group, fill = fill)) + geom_tile( color = "grey70")+ scale_fill_manual(breaks = c(FALSE, TRUE), values = c("white","lightblue2")) +
      geom_text(aes(label = Freq)) + theme_minimal() + theme(legend.position = "none")

# ggsave("plots/SuppFig3_simulated.png", height = 4, width = 5)
```

```{r}
mclust::adjustedRandIndex(sim$clincluster, sim$Group)
```

## Visualisation

```{r}
p1 <- plotPCA(sim, colour_by = "Batch")
p2 <- plotPCA(sim, colour_by = "Group")
p3 <- plotPCA(sim, colour_by = "clincluster")
cowplot::plot_grid(p1,p2,p3, cols = 3)
# ggsave("plots/SuppFig3_splatter_pca1.png", width = 18, height = 5,  dpi = 300)
```

```{r}
mclust::adjustedRandIndex(sim$clincluster, sim$Group)
```



## Technical

```{r}
sessionInfo()
```

R version 3.5.2 (2018-12-20)
Platform: x86_64-apple-darwin15.6.0 (64-bit)
Running under: macOS Mojave 10.14.4

Matrix products: default
BLAS: /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
LAPACK: /Library/Frameworks/R.framework/Versions/3.5/Resources/lib/libRlapack.dylib

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

attached base packages:
 [1] grid      parallel  stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] modes_0.7.0                 ROCR_1.0-7                  gplots_3.0.1.1              KernSmooth_2.23-15          fields_9.6                  maps_3.3.0                 
 [7] spam_2.2-2                  dotCall64_1.0-0             DoubletFinder_2.0.1         Seurat_2.3.4                Matrix_1.2-15               cowplot_0.9.4              
[13] bindrcpp_0.2.2              reshape2_1.4.3              scales_1.0.0                dplyr_0.7.8                 edgeR_3.24.3                limma_3.38.3               
[19] scater_1.10.1               ggplot2_3.1.0               SingleCellExperiment_1.4.1  SummarizedExperiment_1.12.0 DelayedArray_0.8.0          BiocParallel_1.16.5        
[25] matrixStats_0.54.0          Biobase_2.42.0              GenomicRanges_1.34.0        GenomeInfoDb_1.18.1         IRanges_2.16.0              S4Vectors_0.20.1           
[31] BiocGenerics_0.28.0        

loaded via a namespace (and not attached):
  [1] reticulate_1.10          R.utils_2.7.0            tidyselect_0.2.5         htmlwidgets_1.3          trimcluster_0.1-2.1      Rtsne_0.15               devtools_2.0.1          
  [8] munsell_0.5.0            codetools_0.2-16         ica_1.0-2                statmod_1.4.30           scran_1.10.2             umap_0.1.0.3             withr_2.1.2             
 [15] colorspace_1.4-0         knitr_1.21               rstudioapi_0.9.0         robustbase_0.93-3        dtw_1.20-1               gbRd_0.4-11              Rdpack_0.10-1           
 [22] labeling_0.3             lars_1.2                 GenomeInfoDbData_1.2.0   pheatmap_1.0.12          bit64_0.9-7              rhdf5_2.26.2             rprojroot_1.3-2         
 [29] xfun_0.4                 diptest_0.75-7           R6_2.3.0                 ggbeeswarm_0.6.0         locfit_1.5-9.1           hdf5r_1.0.1              flexmix_2.3-14          
 [36] bitops_1.0-6             assertthat_0.2.0         SDMTools_1.1-221         nnet_7.3-12              beeswarm_0.2.3           gtable_0.2.0             npsurv_0.4-0            
 [43] processx_3.2.1           rlang_0.3.1              splines_3.5.2            lazyeval_0.2.1           acepack_1.4.1            checkmate_1.9.1          yaml_2.2.0              
 [50] backports_1.1.3          Hmisc_4.2-0              tools_3.5.2              usethis_1.4.0            RColorBrewer_1.1-2       proxy_0.4-22             dynamicTreeCut_1.63-1   
 [57] sessioninfo_1.1.1        ggridges_0.5.1           kknn_1.3.1               Rcpp_1.0.0               plyr_1.8.4               base64enc_0.1-3          zlibbioc_1.28.0         
 [64] purrr_0.3.0              RCurl_1.95-4.11          ps_1.3.0                 prettyunits_1.0.2        rpart_4.1-13             pbapply_1.4-0            viridis_0.5.1           
 [71] zoo_1.8-4                cluster_2.0.7-1          fs_1.2.6                 magrittr_1.5             data.table_1.12.0        lmtest_0.9-36            RANN_2.6.1              
 [78] mvtnorm_1.0-8            fitdistrplus_1.0-14      pkgload_1.0.2            evaluate_0.13            lsei_1.2-0               mclust_5.4.2             gridExtra_2.3           
 [85] compiler_3.5.2           tibble_2.0.1             crayon_1.3.4             R.oo_1.22.0              htmltools_0.3.6          segmented_0.5-3.0        Formula_1.2-3           
 [92] snow_0.4-3               tidyr_0.8.2              MASS_7.3-51.1            fpc_2.1-11.1             cli_1.0.1                R.methodsS3_1.7.1        gdata_2.18.0            
 [99] metap_1.1                bindr_0.1.1              igraph_1.2.3             pkgconfig_2.0.2          foreign_0.8-71           foreach_1.4.4            vipor_0.4.5             
[106] XVector_0.22.0           bibtex_0.4.2             stringr_1.4.0            callr_3.1.1              digest_0.6.18            tsne_0.1-3               rmarkdown_1.11          
[113] htmlTable_1.13.1         DelayedMatrixStats_1.4.0 curl_3.3                 kernlab_0.9-27           gtools_3.8.1             modeltools_0.2-22        nlme_3.1-137            
[120] jsonlite_1.6             Rhdf5lib_1.4.2           BiocNeighbors_1.0.0      desc_1.2.0               viridisLite_0.3.0        pillar_1.3.1             lattice_0.20-38         
[127] httr_1.4.0               DEoptimR_1.0-8           pkgbuild_1.0.2           survival_2.43-3          glue_1.3.0               remotes_2.0.2            png_0.1-7               
[134] prabclus_2.2-7           iterators_1.0.10         bit_1.1-14               class_7.3-15             stringi_1.2.4            HDF5Array_1.10.1         mixtools_1.1.0          
[141] doSNOW_1.0.16            latticeExtra_0.6-28      caTools_1.17.1.1         memoise_1.1.0            irlba_2.3.3              ape_5.2             
